Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-06 paper/paper.tex
remove distinct mechanisms section of Results (last results section)

\subsection{Distinct characteristics of IFN induction by different viral variants}

%%% begin mutantcomparison figure %%%
\begin{figure}

\centerline{\includegraphics[width=0.85\textwidth]{figures/MutantComparison/p_merged.pdf}}
\caption{
For NS1-deficient infections, higher viral gene expression causes more IFN induction.
{\bf (A)}
Flow cytometry shows that only for viruses with defects in NS1, there is more IFN induction among infected cells that express higher levels of HA protein.
The y-axis shows the ratio of the percent of IFN+ cells in the highest HA-expression quartile relative to the lowest HA-expression quartile.
Points indicate replicates, and lines indicate the mean.
See \FIGSUPP[mutantcomparison]{ha_vs_ifn} for details.
{\bf (B)}
Blocking viral secondary transcription strongly decreases IFN levels in cells infected with NS1stop virus, but not cells infected with virus carrying a deletion in PB1.
A549 cells were infected with the viral stocks at doses that induced equivalent IFN, and ribavirin was added at a concentration (100 $\mu$M) sufficient to block secondary transcription~\citep{Vanderlinden:2016ec,Reuther:2015ef,Scholtissek:1976wg}.
After eight hours, IFNB1 transcript was quantified relative to the housekeeping gene L32 by qPCR. 
The dashed gray line indicates the limit of detection.
\FIGSUPP[mutantcomparison]{cycloheximide} shows that results are similar if secondary transcription is instead blocked with cycloheximide.
}
\label{fig:mutantcomparison}

\figsupp[More detailed data on the flow cytometry analysis in \FIG{mutantcomparison}A.]
{Analysis of the flow cytometry data in \FIGSUPP[validation]{SNPflow} indicates that a higher viral burden is associated with increased IFN-induction in NS1-deficient infections.
We re-analyzed all the infected cells (defining ``infected'' using the same HA-gating scheme in \FIGSUPP[validation]{SNPflow}) for the wild-type virus, the NS1stop virus, and the three amino-acid mutants with significantly increased IFN induction.
For each virus and replicate, we binned the infected cells into HA expression quartiles.
We then calculated the percent of cells that were IFN+ in each quartile (defining ``IFN+'' using the same gating scheme in \FIGSUPP[validation]{SNPflow}).
The plots show the mean HA expression in each quartile versus the percent of cells that are IFN+.
The results clearly indicate that for the NS1stop virus, increased viral burden (more HA) correlates with IFN induction.
A similar trend also holds for the virus with an amino-acid substitution in NS1 (NS1-A122V). 
But for wild type and the two PB1 mutants, there is no clear trend for higher viral burden to correlate with IFN induction.
\FIG{mutantcomparison}A summarizes the data in this figure by simply showing the ratio of the percent IFN+ of the highest HA-expression quartile to the lowest HA-expression quartile for each sample and replicate.}
{\includegraphics[width=0.9\textwidth]{figures/MutantComparison/p_ifn_vs_ha.pdf}}
\label{figsupp:ha_vs_ifn}

\figsupp[Similar to \FIG{mutantcomparison}B, but uses cycloheximide rather than ribavirin.]
{Preventing viral secondary transcription using cycloheximide also decreases IFN induction by the NS1stop virus relative to the PB1 deletion virus, corroborating the result in \FIG{mutantcomparison}B with a different inhibitor.
However, the results are more complex here, since cycloheximide alone increases IFN production~\citep{raj1981analysis, maroteaux1983cycloheximide, killip2014activation}.
Therefore, we see an increase in IFN levels upon cycloheximide treatment for uninfected cells and cells infected with the PB1 deletion virus, but a decrease for cells infected with the NS1stop virus.
This plot shows qPCR results for the same samples as in \FIGSUPP[validation]{cycloheximide}, but with the qPCR for expression of IFNB1 transcripts relative to the housekeeping gene L32.
As described in \FIGSUPP[validation]{cycloheximide}, A549 cells were infected at an MOI of 0.4, such that HA expression was equivalent between viral variants when secondary transcription was blocked, and some of the samples treated with 50 $\mu$g/ml of cycloheximide~\citep[a concentration sufficient to block secondary transcription;][]{killip2014activation}.
After 8 hours, transcript levels were quantified by qPCR.
Points indicate four replicates, and lines indicate the mean.
The dashed line shows the limit of detection, and points below this limit where set to that value.
}
{\includegraphics[width=0.6\textwidth]{figures/MutantComparison/cyclo_qPCR_plot.pdf}}
\label{figsupp:cycloheximide}

\end{figure}
%%% end mutantcomparison figure%%%

The single-cell data reveal differences among IFN+ cells depending on the defect in the virion that infects them.
Specifically, \FIG{mutations}D shows that if a cell is infected by a virion that fails to express NS, then a higher viral transcriptional load correlates with IFN induction---but that no such association is present among infected cells that express NS.
To validate this conclusion, we used flow cytometry to compare the frequency of IFN induction among infected cells that express low or high levels of a viral protein (HA).
\FIG{mutantcomparison}A shows that upon infection with stocks of the NS1stop or NS1-A122V mutant, cells that express high levels of HA are much more likely to be IFN+ than those with low HA expression.
In contrast, if cells are infected with stocks of wild-type or PB1 amino-acid mutant virus, there is only a slight increase in IFN induction among cells with high HA expression.

To determine if higher viral transcriptional burden \emph{causes} IFN induction in NS-deficient infections, we treated infected cells with two different chemicals (ribavirin and cycloheximide) that inhibit secondary viral transcription, and so greatly reduce the level of viral transcripts~\citep{Vanderlinden:2016ec,Reuther:2015ef,Scholtissek:1976wg, killip2014activation}.
As shown in \FIG{mutantcomparison}A and \FIGSUPP[mutantcomparison]{cycloheximide}, both inhibitors markedly reduced IFN induction by cells infected with the NS1stop viral stock, but not by cells infected the virus carrying a deletion in PB1.
This result indicates that high levels of viral transcription promote IFN induction in NS-deficient infections, and suggests that there are distinct mechanisms of IFN induction by different viral variants---perhaps explaining why prior studies have not reached a consensus on the process by which influenza induces IFN~\citep{rehwinkel2010rig, weber2013incoming, killip2014activation}.

@@

For the qPCR in \FIGSUPP[validation]{delqPCR} and \FIGSUPP[mutantcomparison]{cycloheximide}, A549 cells were seeded at a density of $10^4$ cells/well in a 96-well plate in D10 media 24 hours prior to infection, with four independent wells seeded per experimental treatment. 
Immediately prior to infection D10 media was removed and replaced with influenza growth media and infected with the indicated influenza strains at a MOI of 0.4 based on TCID50 in MDCK-SIAT1 cells.
For the cells with cycloheximide added to block protein expression (and hence secondary transcription), cycloheximide was added to a final concentration of 50 $\mu$g/ml~\citep[a concentration sufficient to block secondary transcription;][]{killip2014activation} at the time of infection.
After 8 hours, mRNA was harvested using the CellAmp Direct RNA Prep Kit for RT-PCR, reverse-transcribed using an oligo-dT primer, and qPCR was performed as described above.

For the qPCR in \FIG{mutantcomparison}B, A549 cells were infected with the NS1stop virus at a MOI of 0.1 (based on TCID50) in MDCK-SIAT1 and with the PB1del385to2163 virus at a dose that induced equivalent IFN in the absence of ribavirin.
The ribavirin was added to cells at the time of infection to a final concentration of 100$\mu$M, which is sufficient to block secondary transcription~\citep{Vanderlinden:2016ec,Reuther:2015ef,Scholtissek:1976wg}.
All other steps were performed as for the cycloheximide experiments described immediately above..

>>>>

@@

For the qPCR in \FIGSUPP[validation]{delqPCR}, A549 cells were seeded at a density of $10^4$ cells/well in a 96-well plate in D10 media 24 hours prior to infection, with four independent wells seeded per experimental treatment. 
Immediately prior to infection D10 media was removed and replaced with influenza growth media and infected with the indicated influenza strains at a MOI of 0.4 based on TCID50 in MDCK-SIAT1 cells.
For the cells with cycloheximide added to block protein expression (and hence secondary transcription), cycloheximide was added to a final concentration of 50 $\mu$g/ml~\citep[a concentration sufficient to block secondary transcription;][]{killip2014activation} at the time of infection.
After 8 hours, mRNA was harvested using the CellAmp Direct RNA Prep Kit for RT-PCR, reverse-transcribed using an oligo-dT primer, and qPCR was performed as described above.
